---
title: "Clustering Wine Dataset"
author: "Anushree Shivarudrappa"
date: "June 9, 2016"
output: pdf_document
---

## This mini-project is based on the K-Means exercise from 'R in Action'
Go here for the original blog post and solutions
http://www.r-bloggers.com/k-means-clustering-from-r-in-action/

## 1. Exercise 0: Install these packages if you don't have them already

**install.packages(c("cluster", "rattle","NbClust"))**
```{r, warning=FALSE, message=FALSE}
library(cluster)
library(rattle)
library(NbClust)
```
  
**Now load the data and look at the first few rows**
```{r}
data(wine, package="rattle")
head(wine)
```
## 2. Exercise 1: Remove the first column from the data and scale it using the scale() function
```{r}
df_wine <- scale(wine[-1])
summary(df_wine)
```
  
**Now we'd like to cluster the data using K-Means.**  
**How do we decide how many clusters to use if you don't know that already?**  
**We'll try two methods.** 

##Method 1:
A plot of the total within-groups sums of squares against the number of clusters in a K-means solution can be helpful. A bend in the graph can suggest the appropriate number of clusters. 

```{r}
wssplot <- function(data, nc=15, seed=1234){
	              wss <- (nrow(data)-1)*sum(apply(data,2,var))
               	      for (i in 2:nc){
		        set.seed(seed)
	                wss[i] <- sum(kmeans(data, centers=i)$withinss)}
	                
		      plot(1:nc, wss, type="b", xlab="Number of Clusters",
	                        ylab="Within groups sum of squares")
	   }

wssplot(df_wine)
```

## 3. Exercise 2:  
**How many clusters does this method suggest?**  
Method suggest cluster count as 3 i.e k=3.  
  
**Why does this method work? What's the intuition behind it?**  
Sum of square error (SSE) plot is used to determine appropriate k value. Plot indicates that there is distinct drop when moving from 1 to 3 clusters. After 3 we can observe decrease in drop off, this suggest a 3-cluster as solution. If we look at the original data it also contains 3 classes.  

##Method 2:
Use the NbClust library, which runs many experiments and gives a distribution of potential number of clusters.

```{r, fig.show='hide', results='hide', message=FALSE, warning=FALSE}
library(NbClust)
set.seed(1234)
nc <- NbClust(df_wine, min.nc=2, max.nc=15, method="kmeans")
```
```{r, fig.width=9}
barplot(table(nc$Best.n[1,]),
	          xlab="Numer of Clusters", ylab="Number of Criteria",
		            main="Number of Clusters Chosen by 26 Criteria")

```
  
## 4. Exercise 3: How many clusters does this method suggest?  
By Looking into graph we can say best number of cluster is 3 i.e k=3          

## 5. Exercise 4: Once you've picked the number of clusters, run k-means using this number of clusters. Output the result of calling kmeans() into a variable fit.km

```{r}
fit.km <- kmeans( df_wine, 3 )
str(fit.km)
```
Now we want to evaluate how well this clustering does.

## 6. Exercise 5: 
**Using the table() function, show how the clusters in fit.km$clusters
compares to the actual wine types in wine$Type. Would you consider this a good clustering?**
```{r}
table(fit.km$cluster)
table(wine$Type)
table_clust_wine <- table(wine$Type, fit.km$cluster)
table_clust_wine 
```
Its a confusion matrix. To find "fit.km" is a good cluster or not, using an adjusted Rank index provided by the flexclust package.
```{r, warning=FALSE, message=FALSE}
library(flexclust)
randIndex(table_clust_wine)
```
The adjusted Rand index provides a measure of the agreement between two partitions, adjusted for chance. It ranges from -1 (no agreement) to 1 (perfect agreement). Agreement between the wine varietal type and the cluster solution is 0.9. So its a good clustering

## 7. Exercise 6:
**Visualize these clusters using  function clusplot() from the cluster library**
**Would you consider this a good clustering?**
```{r}
clusplot( df_wine, fit.km$cluster, color = T, shade = T, 
          main='2D representation of the Cluster solution')
```